---
title: "Week 9 Assignment"
author: "James Naval"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(keyring)
library(tidyverse)
library(jsonlite)
library(DT)
```

## Introduction {.tabset}

In this assignment we have been asked to use the [New York Times Developer API](https://developer.nytimes.com/) to obtain information and pull it into a R data frame. 

### Using the MOST Popular API

For the first example I will use the MOST POPULAR API, to load the most shared articles by facebook.

```{r get-data-using-Most-Popular-api}

url <- paste("https://api.nytimes.com/svc/mostpopular/v2/shared/1/facebook.json?api-key=8Tuw2G2y7TwdrHs4sejvtQWjJ2E844o5")

rawDataFromMostPopularApi <- fromJSON(url,flatten = T)
resultsDataFramemostpopular <- rawDataFromMostPopularApi$results

resultsDataFramemostpopular %>% 
  select(title, section, source,  abstract, type, byline, adx_keywords, published_date,updated) %>% 
  datatable()

```

### Display Results Graphically for Most Popular API

I will now make a bar graph showing the number of popular articles per section.
```{r ggplot-section}
resultsDataFramemostpopular %>% ggplot(aes(x=section, fill = section)) + geom_bar()
```

### Using the Top Stories API per Science

For the second example I will use the Top Stories API, to load the top science stories.

```{r get-data-using-top-stories-api-science}

url <- paste("https://api.nytimes.com/svc/topstories/v2/science.json?api-key=8Tuw2G2y7TwdrHs4sejvtQWjJ2E844o5")

rawDataFromTopStoriesApi <- fromJSON(url,flatten = T)
resultsDataFrameTopStories <- rawDataFromTopStoriesApi$results

resultsDataFrameTopStories %>% 
  select( title,section, short_url, abstract) %>% 
  datatable()

```

### Display Results Graphically for Top Stories API per science

I will now make a bar graph showing the number of articles per section.
```{r ggplot-section-science}
resultsDataFrameTopStories %>% ggplot(aes(x=section, fill = section)) + geom_bar()
```

### Using the Top Stories API per Technology

For the third example I will use the Top Stories API, to load the top technology stories.

```{r get-data-using-top-stories-api-technology}

url <- paste("https://api.nytimes.com/svc/topstories/v2/technology.json?api-key=8Tuw2G2y7TwdrHs4sejvtQWjJ2E844o5")

rawDataFromTopStoriesApiTechnology <- fromJSON(url,flatten = T)
resultsDataFrameTopStoriesTechnology <- rawDataFromTopStoriesApiTechnology$results

resultsDataFrameTopStories %>% 
  select( title,section, short_url, abstract) %>% 
  datatable()

```

### Display Results Graphically for Top Stories API per technology

I will now make a bar graph showing the number of articles per section.
```{r ggplot-section-technology}
resultsDataFrameTopStoriesTechnology %>% ggplot(aes(x=section, fill = section)) + geom_bar()
```

## Conclusion

As can be seen, it is quite useful and simple to access data using an API that returns JSON, I used three different API's and the results of these API are display above.