---
title: "DATA607 Week 2 Assignment"
author: "James Naval"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RCurl)
library(RODBC)
library(odbc)
library(DBI)
library(RSQLite)
library(elo)
```

### Introduction

For this assignment we were suppose to conduct a survey asking at least 5 people to rate the last 10 movies on a scale of 1-5. For this survey I asked 10 people which consist of some co-workers and friends to rate they most recent movies. After everyone complete the survey I record the result and upload. The results for the survey is available for download from either google.com or github.com using the link below:

## Google: <https://docs.google.com/spreadsheets/d/13nVSyE9lhiyPPZTH_HYi9FNYIfYapg3x/edit#gid=45569196>
## Github: <https://raw.githubusercontent.com/jnaval88/DATA607/main/Most_recent_Movies_Rating.csv.csv>

The movies that were using for the survey were: Orion and the Dark, The Beekeeper, Night Swim, Lift, Justice League: Crisis on infinite Earth Part One, The Bricklayer, Badland Hunters, Mean Girls, Role Play and The Tiger's Apprentice. For this assignment I will use the R build in SQL database and store the table into memory.

### Github Data Importing  into R

```{r ImportingCSVfilefromGithub}

fileUrl <- getURL('https://raw.githubusercontent.com/jnaval88/DATA607/main/Most_recent_Movies_Rating.csv.csv')
Most_recent_Movies_Rating <- read.csv(text = fileUrl)
Most_recent_Movies_Rating

colnames(Most_recent_Movies_Rating)

```

### Creating SQL Database using Build SQL in R

```{r loadingCSVfile to  SQLDatabase}

con <- dbConnect(SQLite(), "memory")
copy_to(con,Most_recent_Movies_Rating1)
Most_recent_Movies_Rating
```

## Table into SQL using GetQuery

```{r loading table from SQL QUERY}

dbGetQuery(con,'
SELECT *
FROM Most_recent_Movies_Rating1
           ')

```

## Getting Survey score in Descending Order

```{r Display SCORE Rating SQL}

dbGetQuery(con, '
SELECT SCORE, COUNT(SCORE)
FROM Most_recent_Movies_Rating1
GROUP BY SCORE
ORDER BY COUNT(SCORE) DESC
           ')
```

### Display two Column from SQL using GetQuery

```{r joining column}

dbGetQuery(con, '
SELECT Most_recent_Movies_Rating1.SCORE, Most_recent_Movies_Rating1.TITLE
FROM Most_recent_Movies_Rating1
JOIN (SELECT SCORE, TITLE as opponent from Most_recent_Movies_Rating1) B
ON Most_recent_Movies_Rating1.SCORE = b.SCORE
ORDER BY Most_recent_Movies_Rating1.SCORE DESC
           
           ')

```

```{sql, connection = con, output.var = "df"}

SELECT Most_recent_Movies_Rating1.SCORE, Most_recent_Movies_Rating1.TITLE
FROM Most_recent_Movies_Rating1
JOIN (SELECT SCORE, TITLE as opponent from Most_recent_Movies_Rating1) B
ON Most_recent_Movies_Rating1.SCORE = b.SCORE
ORDER BY Most_recent_Movies_Rating1.SCORE DESC

```

### Creative SQL Query from dplyr

```{r DPLYRSQL}

Most_recent_Movies_Rating1.db = tbl(con, "Most_recent_Movies_Rating1")

Most_recent_Movies_Rating1_query <- Most_recent_Movies_Rating1.db %>%
select(SCORE, TITLE, MAIN_GENRE) %>%
left_join(Most_recent_Movies_Rating1.db %>% select(MAIN_GENRE, SCORE = TITLE))

Most_recent_Movies_Rating1_query %>% show_query()

```
### SQL QUERY from Dplyr

```{sql, connection = con}

SELECT
  `Most_recent_Movies_Rating1_LHS`.`SCORE` AS `SCORE`,
  `Most_recent_Movies_Rating1_LHS`.`TITLE` AS `TITLE`,
  `Most_recent_Movies_Rating1_LHS`.`MAIN_GENRE` AS `MAIN_GENRE`
FROM `Most_recent_Movies_Rating1` AS `Most_recent_Movies_Rating1_LHS`
LEFT JOIN `Most_recent_Movies_Rating1` AS `Most_recent_Movies_Rating1_RHS`
  ON (
    `Most_recent_Movies_Rating1_LHS`.`SCORE` = `Most_recent_Movies_Rating1_RHS`.`TITLE` AND
    `Most_recent_Movies_Rating1_LHS`.`MAIN_GENRE` = `Most_recent_Movies_Rating1_RHS`.`MAIN_GENRE`
  )

```
### Create Table

```{sql, connection = con, "memory"}
CREATE TABLE Recent_Movies_Rating(
     TITLE CHAR(50) PRIMARY KEY( TITLE) NOT NULL,
   RELEASE_YEAR CHAR(50) NOT NULL,
   SCORE INT NOT NULL ,
   NUMBER_OF_VOTES INT NOT NULL,
   DURATION CHAR(50) NOT NULL,
  MAIN_GENRE CHAR(50) NOT NULL,
  MAIN_PRODUCTION CHAR(50) NOT NULL,
);
```


### Conclusion

For this lab I took data from a survey that I conducted and inserted it in R build in  SQL database. Base on the rating we can that Orion and the Dark have the highest rating followed by the BeeKeeper then The Tiger's Apprentice.


